---
title: "Take-home_Ex03"
author: "CHJ1"
---

Installing Packages

```{r}
pacman::p_load(httr, tidyverse, sf, tmap, jsonlite, progress, dplyr, olsrr, corrplot, ggpubr, sfdep, GWmodel, gtsummary, performance, RColorBrewer)
```

Geocoding HDB Resale Data

Select Data of Interest

```{r}
HDBresale <- read_csv(
  "data/aspatial/HDBresale.csv") %>%
  filter(flat_type == "5 ROOM") %>%
  filter(month >= "2025-01" & month <= "2025-09")
```

```{r}
geocode <- function(block, streetname) {
  base_url <- "https://onemap.gov.sg/api/common/elastic/search"
  address <- paste(block, streetname, sep = " ")
  query <- list("searchVal" = address, 
                "returnGeom" = "Y",
                "getAddrDetails" = "N",
                "pageNum" = "1")
  
  res <- GET(base_url, query = query)
  restext<-content(res, as="text")
  
  output <- fromJSON(restext)  %>% 
    as.data.frame %>%
    select(results.LATITUDE, results.LONGITUDE)

  return(output)
}
```

```{r}
HDBresale$LATITUDE <- 0
HDBresale$LONGITUDE <- 0

for (i in 1:nrow(HDBresale)){
  temp_output <- geocode(HDBresale[i, 4], HDBresale[i, 5])
  
  HDBresale$LATITUDE[i] <- temp_output$results.LATITUDE
  HDBresale$LONGITUDE[i] <- temp_output$results.LONGITUDE
}
```

Converting HDBresale to an sf object

```{r}
HDBresale_sf <- HDBresale %>%
  st_as_sf(
    coords = c("LONGITUDE", "LATITUDE"),
    crs=4326) %>%
  st_transform(crs = 3414)
```

Import Preschool Location data

```{r}
preschool = st_read(
  "data/aspatial/PreSchoolsLocation.kml") %>%
  st_transform(crs= 3414)
```

Counting number of preschools

```{r}
prischool_within_1000m <- st_is_within_distance(
  HDBresale_sf, preschool, dist = 1000)

HDBresale_sf <- HDBresale_sf %>%
  mutate(prischool_within_1000m = map_int(prischool_within_1000m, length))
```

Import Childcare Location data

```{r}
childcare = st_read(
  "data/aspatial/ChildCareServices.kml") %>%
  st_transform(crs= 3414)
```

Counting number of childcare

```{r}
childcare_within_350m <- st_is_within_distance(
  HDBresale_sf, childcare, dist = 350)

HDBresale_sf <- HDBresale_sf %>%
  mutate(childcare_within_350m = map_int(childcare_within_350m, length))
```

Import kindergartens Location data

```{r}
kindergartens = st_read(
  "data/aspatial/Kindergartens.kml") %>%
  st_transform(crs= 3414)
```

Counting number of kindergarten

```{r}
kindergarten_within_350m <- st_is_within_distance(
  HDBresale_sf, kindergartens, dist = 350)

HDBresale_sf <- HDBresale_sf %>%
  mutate(kindergarten_within_350m = map_int(kindergarten_within_350m, length))
```

Import bus stop Location data

```{r}
busstop = st_read(dsn = "data/aspatial/BusStopLocation_Aug2025/",
                  layer = "BusStop") %>%
  st_transform(crs = 3414)
```

Counting number of busstop

```{r}
busstop_within_350m <- st_is_within_distance(
  HDBresale_sf, busstop, dist = 350)

HDBresale_sf <- HDBresale_sf %>%
  mutate(busstop_within_350m = map_int(busstop_within_350m, length))
```

Import eldercare service Location data

```{r}
eldercare = st_read(
  "data/aspatial/EldercareServicesKML.kml") %>%
  st_transform(crs= 3414)
```

compute proximity to eldercare service

```{r}
HDBresale_sf <- HDBresale_sf %>%
  mutate(
    PROX_eldercare = as.numeric(
      st_distance(geometry,
                  eldercare[st_nearest_feature(
                    geometry, eldercare), ], 
                  by_element = TRUE)
      )
    )
```

Import lrt mrt Location data

```{r}
lrtmrt = st_read(
  "data/aspatial/LTAMRTStationExitKML.kml") %>%
  st_transform(crs= 3414)
```

compute proximity to lrt and mrt

```{r}
HDBresale_sf <- HDBresale_sf %>%
  mutate(
    PROX_lrtmrt = as.numeric(
      st_distance(geometry,
                  lrtmrt[st_nearest_feature(
                    geometry, lrtmrt), ], 
                  by_element = TRUE)
      )
    )
```

Import hawker centre Location data

```{r}
hawker = st_read(
  "data/aspatial/HawkerCentresKML.kml") %>%
  st_transform(crs= 3414)
```

compute proximity to hawker centre

```{r}
HDBresale_sf <- HDBresale_sf %>%
  mutate(
    PROX_hawker = as.numeric(
      st_distance(geometry,
                  hawker[st_nearest_feature(
                    geometry, hawker), ], 
                  by_element = TRUE)
      )
    )
```

Import market and food centre Location data

```{r}
marketfoodcentre = st_read(
  "data/aspatial/NEAMarketandFoodCentreKML.kml") %>%
  st_transform(crs= 3414)
```

compute proximity to market and food centre

```{r}
HDBresale_sf <- HDBresale_sf %>%
  mutate(
    PROX_marketfoodcentre = as.numeric(
      st_distance(geometry,
                  marketfoodcentre[st_nearest_feature(
                    geometry, marketfoodcentre), ], 
                  by_element = TRUE)
      )
    )
```

Import supermarket location data

```{r}
supermarket <- read_csv("data/aspatial/ListofSupermarketLicences.csv")
```

Geocode supermarket postal codes

```{r}
postcodes <- unique(supermarket$postal_code)

url <- "https://onemap.gov.sg/api/common/elastic/search"

found <- data.frame()
not_found <- data.frame(postcode = character())

for (pc in postcodes) {
  query <- list(
    searchVal = pc,
    returnGeom = "Y",
    getAddrDetails = "Y",
    pageNum = "1"
  )
  
  res <- GET(url, query = query)
  json <- content(res)
  
  if (json$found != 0) {
    df <- as.data.frame(json$results, stringsAsFactors = FALSE)
    df$input_postcode <- pc
    found <- bind_rows(found, df)
  } else {
    not_found <- bind_rows(not_found, data.frame(postcode = pc))
  }
}
```

Appending location information to supermarket

```{r}
supermarket = supermarket %>%
  left_join(found, 
            by = c('postal_code' = 'POSTAL'))
```

Convert supermarket to sf object

```{r}
supermarket_clean <- supermarket %>%
  filter(!is.na(X) & !is.na(Y))

supermarket_sf <- st_as_sf(supermarket_clean, 
                         coords = c("X","Y"),
                         crs=3414) 
```

Compute proximity to supermarket

```{r}
HDBresale_sf <- HDBresale_sf %>%
  mutate(
    PROX_supermarket = as.numeric(
      st_distance(geometry,
                  supermarket_sf[st_nearest_feature(geometry, supermarket_sf), ],
                  by_element = TRUE)
    )
  )
```

Import park location data

```{r}
park <- st_read("data/aspatial/ParkFacilities.geojson") %>%
  st_transform(crs = 3414) 
```

Compute proximity to park

```{r}
HDBresale_sf <- HDBresale_sf %>%
  mutate(
    PROX_Park = as.numeric(
      st_distance(geometry, 
                  park[st_nearest_feature(geometry, park), ],
                  by_element = TRUE)
  ))
```

Using raffles place mrt coordinates as cbd

```{r}
cbd <- st_point(c(103.851959, 1.285400)) 

cbd_sf <- st_sfc(cbd, crs = 4326) %>%
  st_transform(crs = 3414)
```

Compute proximity to cbd

```{r}
HDBresale_sf <- HDBresale_sf %>%
  mutate(
    PROX_CBD = as.numeric(st_distance(geometry, cbd_sf))
  )
```

Compute age of unit

```{r}
HDBresale_sf <- HDBresale_sf %>%
  mutate(age_of_unit = 2025 - lease_commence_date)
```

Compute remaining lease

```{r}
HDBresale_sf <- HDBresale_sf %>%
  mutate(remaininglease = 99 - age_of_unit)
```

Changing storey to continuous value

```{r}
storey_map <- c(
  "01 TO 03" = 1,
  "04 TO 06" = 2,
  "07 TO 09" = 3,
  "10 TO 12" = 4,
  "13 TO 15" = 5,
  "16 TO 18" = 6,
  "19 TO 21" = 7,
  "22 TO 24" = 8,
  "25 TO 27" = 9,
  "28 TO 30" = 10,
  "31 TO 33" = 11,
  "34 TO 36" = 12,
  "37 TO 39" = 13,
  "40 TO 42" = 14,
  "43 TO 45" = 15,
  "46 TO 48" = 16,
  "49 TO 51" = 17
)

HDBresale_sf <- HDBresale_sf %>%
  mutate(floor_level = storey_map[storey_range])
```

Importing geospatial data

```{r}
mpsz = st_read(dsn = "data/geospatial", 
               layer = "MP14_SUBZONE_WEB_PL")
```

Updating CRS information

```{r}
mpsz_svy21 <- st_transform(mpsz, 3414)
```

```{r}
st_crs(mpsz_svy21)
```

```{r}
st_bbox(mpsz_svy21)
```

Exploratory Data Analysis (EDA)

```{r}
ggplot(data=HDBresale_sf, 
       aes(x=`resale_price`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")
```

The figure above reveals a right skewed distribution. This means that more 5 room HDB units were transacted between 600k to 900k.

Logging resale price

```{r}
HDBresale_sf <- HDBresale_sf %>%
  mutate(`LOG_RESALE_PRICE` = log(resale_price))
```

```{r}
ggplot(data=HDBresale_sf, 
       aes(x=`LOG_RESALE_PRICE`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")
```

Multiple Histogram Plots distribution of variables

```{r}
colnames(HDBresale_sf)
```

```{r}
floor_area_sqm <- ggplot(data=HDBresale_sf, 
                               aes(x= `floor_area_sqm`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

prischool_within_1000m <- ggplot(data=HDBresale_sf, 
                               aes(x= `prischool_within_1000m`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

childcare_within_350m <- ggplot(data=HDBresale_sf, 
                               aes(x= `childcare_within_350m`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

kindergarten_within_350m <- ggplot(data=HDBresale_sf, 
                               aes(x= `kindergarten_within_350m`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

busstop_within_350m <- ggplot(data=HDBresale_sf, 
                               aes(x= `busstop_within_350m`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

PROX_eldercare <- ggplot(data=HDBresale_sf, 
                               aes(x= `PROX_eldercare`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

PROX_lrtmrt <- ggplot(data=HDBresale_sf, 
                               aes(x= `PROX_lrtmrt`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

PROX_hawker <- ggplot(data=HDBresale_sf, 
                               aes(x= `PROX_hawker`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

PROX_marketfoodcentre <- ggplot(data=HDBresale_sf, 
                               aes(x= `PROX_marketfoodcentre`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

PROX_supermarket <- ggplot(data=HDBresale_sf, 
                               aes(x= `PROX_supermarket`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

PROX_Park <- ggplot(data=HDBresale_sf, 
                               aes(x= `PROX_Park`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

PROX_CBD <- ggplot(data=HDBresale_sf, 
                               aes(x= `PROX_CBD`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

age_of_unit <- ggplot(data=HDBresale_sf, 
                               aes(x= `age_of_unit`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

remaininglease <- ggplot(data=HDBresale_sf, 
                               aes(x= `remaininglease`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

floor_level <- ggplot(data=HDBresale_sf, 
                               aes(x= `floor_level`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

ggarrange(floor_area_sqm, prischool_within_1000m, childcare_within_350m, kindergarten_within_350m, busstop_within_350m, PROX_eldercare, PROX_lrtmrt, PROX_hawker, PROX_marketfoodcentre, PROX_supermarket, PROX_Park, PROX_CBD, age_of_unit, remaininglease, floor_level, ncol = 3, nrow = 5)
```

Geospatial distribution of 5 room HDB resale prices in Singapore

```{r}
tmap_mode("view")

tm_shape(mpsz_svy21)+
  tm_polygons() +
tm_shape(HDBresale_sf) +  
  tm_dots(fill = "resale_price",
          fill_alpha = 0.6,
          size = 0.5,
          fill.scale = tm_scale_intervals(
            style="quantile")) +
  tm_view(set_zoom_limits = c(11,14))
  
tmap_mode("plot")
```

The above reveals that higher-priced transactions (darker blues) are concentrated in central and southern parts of Singapore, likely reflecting better accessibility and amenities, while lower-priced transactions (lighter blues) tend to occur in northern and western areas, highlighting how property value varies significantly by location and underscoring the spatial impact on resale pricing.

Multiple Linear Regression Method

Visualising the relationships of the independent variables using correlation matrix

```{r}
cols_to_use <- c(
  "floor_area_sqm", "prischool_within_1000m", "childcare_within_350m", "kindergarten_within_350m", 
  "busstop_within_350m", "PROX_eldercare", "PROX_lrtmrt", "PROX_hawker", "PROX_marketfoodcentre",
  "PROX_supermarket", "PROX_Park", "PROX_CBD", "age_of_unit", "remaininglease", "floor_level"
)

selected_data <- HDBresale_sf[, cols_to_use]

selected_data_numeric <- as.data.frame(
  lapply(selected_data, function(x) as.numeric(as.character(x)))
)

corrplot(cor(selected_data_numeric[, 1:15]), 
         diag = FALSE, 
         order = "AOE",
         tl.pos = "td", 
         tl.cex = 0.5, 
         method = "number", 
         type = "upper")
```

remaininglease is highly correlated to age_of_unit. Thus, exclude remaininglease from the model.

Building a hedonic pricing model using multiple linear regression method

```{r}
five_rm_mlr <- lm(
  formula = resale_price ~ floor_area_sqm + prischool_within_1000m + childcare_within_350m + kindergarten_within_350m + busstop_within_350m + PROX_eldercare + PROX_lrtmrt + PROX_hawker + PROX_marketfoodcentre + PROX_supermarket + PROX_Park + PROX_CBD + age_of_unit + floor_level,
  data=HDBresale_sf)
summary(five_rm_mlr)
```

Remove independent variables that are not statistically significant.

```{r}
five_rm_mlr <- lm(
  formula = resale_price ~ floor_area_sqm + prischool_within_1000m + kindergarten_within_350m + busstop_within_350m + PROX_lrtmrt + PROX_hawker + PROX_marketfoodcentre + PROX_supermarket + PROX_Park + PROX_CBD + age_of_unit + floor_level,
  data=HDBresale_sf)
summary(five_rm_mlr)
```

The output above reveals that all explanatory variables are statistically significant at 95% confident level.

Create formatted regression report

```{r}
tbl_regression(five_rm_mlr, intercept = TRUE)
```

Creating scatter plots with trendline for each statistically significant variable against resale price

```{r}
variables <- c(
  "floor_area_sqm",
  "prischool_within_1000m",
  "kindergarten_within_350m",
  "busstop_within_350m",
  "PROX_lrtmrt",
  "PROX_hawker",
  "PROX_marketfoodcentre",
  "PROX_supermarket",
  "PROX_Park",
  "PROX_CBD",
  "age_of_unit",
  "floor_level"
)

for (v in variables) {
  p <- ggplot(HDBresale_sf, aes_string(x = v, y = "resale_price")) +
    geom_point(alpha = 0.4, color = "blue") +
    geom_smooth(method = "lm", se = FALSE, color = "red", linewidth = 1) +
    labs(title = paste("Scatter plot of", v, "vs resale_price"))
  
  print(p)
}
```

Based on the scatterplots with trendlines for each variable against resale price, only variables whose trendlines are noticeably steeper are highlighted, as these demonstrate a stronger relationship with resale price and indicate that changes in these variables have a more substantial impact on resale values compared to predictors with flatter or negligible slopes.

1. prischool_within_1000m increases, resale price decreases
2. PROX_hawker increases, resale price decreases
3. PROX_marketfoodcentre increases, resale price decreases
4. PROX_CBD increases, resale price decreases
5. age_of_unit increases, resale price decreases
6. floor_level increases, resale price increases

There is stronger demand for 5 room HDB units that offer convenience by being located nearer to food places, closer to the CBD (workplaces), as well as for newer units and those on higher floors. This preference arises because proximity to essential amenities and workplaces reduces travel time and daily hassle, newer units typically offer better conditions and longer lease terms, and higher floors often provide better views, privacy, and less noise, all of which enhance the living experience and attract buyers willing to pay a premium.

There is lower demand for units near primary schools, possibly due to noise from school activities, making these locations less desirable.

Multicollinearity test

```{r}
check_collinearity(five_rm_mlr)
```

Test for Non-Linearity

```{r}
check_model(five_rm_mlr,
            check = "linearity")
```

The green line is not flat and horizontal; instead, it shows a curved pattern.

This indicates the presence of non-linearity in the data: the relationship between one or more variables and the outcome is not fully captured by a linear model.

The linearity assumption of the multiple linear regression model is violated. Predictions will be biased for high and low fitted values.

Test for Normality Assumption

```{r}
check_normality(five_rm_mlr)
```

The p-value of the normality assumption test is less than alpha value of 0.05. Hence, we reject the normality assumption at 95% confident level.

```{r}
plot(check_normality(five_rm_mlr), 
     type = "qq")
```

Q-Q plot above shows that majority of the data points are along the zero line.

Testing for Spatial Autocorrelation

```{r}
mlr.output <- as.data.frame(five_rm_mlr$residuals)
```

```{r}
HDBresale_sf <- cbind(HDBresale_sf, 
                        five_rm_mlr$residuals) %>%
rename(`MLR_RES` = `five_rm_mlr.residuals`)
```

```{r}
tmap_mode("view")

tm_shape(mpsz_svy21) +
  tm_polygons(fill_alpha = 0.4) +
tm_shape(HDBresale_sf) +  
  tm_dots(
    fill = "MLR_RES",
    size = 0.7,
    col = "black",
    fill.scale = tm_scale(
      n = 10,
      values = rev(
        brewer.pal(11, "RdBu")),
      style = "quantile",
      midpoint = NA),
    fill.legend = tm_legend(title = "Residuals")
  ) + 
  tm_title("LM Residuals (Quantile Classification)") +
  tm_layout(legend.outside = TRUE) +
  tm_view(set_zoom_limits = c(11,14))

tmap_mode("plot")
```

The plot reveals that there is sign of spatial autocorrelation. To proof that our observation is true, Global Moran's I test will be performed.

```{r}
HDBresale_sf <- HDBresale_sf %>% 
  mutate(
    nb = st_dist_band(
      st_geometry(geometry), 
                 upper = 1500),
    wt = st_weights(
      nb, style = "W"),
    .before = 1) 
```

```{r}
set.seed(1234)
global_moran_perm(
  HDBresale_sf$MLR_RES, 
  nb = HDBresale_sf$nb, 
  wt = HDBresale_sf$wt,
  alternative = "two.sided", 
  nsim = 499)
```

The Global Moran’s I test for residual spatial autocorrelation shows that it’s p-value is less than the alpha value of 0.05. Hence, we will reject the null hypothesis that the residuals are randomly distributed.

Since the Observed Global Moran I = 0.3354 which is greater than 0, we can infer than the residuals resemble cluster distribution.

Building Adaptive Bandwidth GWR Model

Computing the adaptive bandwidth

```{r}
bw.adaptive <- bw.gwr(
  formula = resale_price ~ floor_area_sqm + prischool_within_1000m + kindergarten_within_350m + busstop_within_350m + PROX_lrtmrt + PROX_hawker + PROX_marketfoodcentre + PROX_supermarket + PROX_Park + PROX_CBD + age_of_unit + floor_level,
  data=HDBresale_sf, 
  approach="CV", 
  kernel="gaussian", 
  adaptive=TRUE, 
  longlat=FALSE)
```

The result shows that 36 is the recommended data points to be used.

Constructing the adaptive bandwidth gwr model

```{r}
gwr.adaptive <- gwr.basic(
  formula = resale_price ~ floor_area_sqm + prischool_within_1000m + kindergarten_within_350m + busstop_within_350m + PROX_lrtmrt + PROX_hawker + PROX_marketfoodcentre + PROX_supermarket + PROX_Park + PROX_CBD + age_of_unit + floor_level,
  data=HDBresale_sf, 
  bw=bw.adaptive, 
  kernel = 'gaussian', 
  adaptive=TRUE, 
  longlat = FALSE)
```

```{r}
gwr.adaptive
```

The report shows that the AICc of the adaptive distance gwr is 111469.9 

Visualising GWR Output

Converting SDF into sf data.frame

```{r}
HDBresale_sf.adaptive <-
  st_as_sf(gwr.adaptive$SDF) %>%
  st_transform(crs=3414)
```

```{r}
glimpse(HDBresale_sf.adaptive)
```

```{r}
summary(gwr.adaptive$SDF$yhat)
```

Visualising local R2

```{r}
tmap_mode("view")

tm_shape(mpsz_svy21)+
  tm_polygons(fill_alpha = 0.1) +
tm_shape(HDBresale_sf.adaptive) +  
  tm_dots(col = "Local_R2",
          border.col = "gray60",
          border.lwd = 1) +
  tm_view(set.zoom.limits = c(11,14))

tmap_mode("plot")
```

Conclusion

The analysis shows that key statistically significant factors influencing 5-room HDB resale prices include floor area, accessibility to preschools and kindergartens, proximity to bus stops, LRT/MRT stations, hawker centres, market food centres, supermarkets, parks, and the CBD, as well as the age of the unit and floor level. Scatterplots with trendlines reveal that variables such as proximity to hawker centres, market food centres, and the CBD, newer units, and higher floor levels have stronger positive impacts on resale price, demonstrating higher demand for convenience, better living conditions, and views. 

Conversely, increased proximity to primary schools correlates with lower resale prices, likely due to noise concerns. Overall, buyers value units that offer accessibility to amenities and workplaces, newer units for better condition and longer lease, and higher floors for enhanced privacy and noise reduction, driving premiums in these areas.



