---
title: "Take-home_Ex03"
author: "CHJ1"
---

Installing Packages

```{r}
pacman::p_load(httr, tidyverse, sf, tmap, 
               jsonlite, progress, dplyr)
```

Geocoding HDB Resale Data

Select Data of Interest

```{r}
HDBresale <- read_csv(
  "data/aspatial/HDBresale.csv") %>%
  filter(flat_type == "5 ROOM") %>%
  filter(month >= "2025-01" & month <= "2025-09")
```

```{r}
geocode <- function(block, streetname) {
  base_url <- "https://onemap.gov.sg/api/common/elastic/search"
  address <- paste(block, streetname, sep = " ")
  query <- list("searchVal" = address, 
                "returnGeom" = "Y",
                "getAddrDetails" = "N",
                "pageNum" = "1")
  
  res <- GET(base_url, query = query)
  restext<-content(res, as="text")
  
  output <- fromJSON(restext)  %>% 
    as.data.frame %>%
    select(results.LATITUDE, results.LONGITUDE)

  return(output)
}
```

```{r}
HDBresale$LATITUDE <- 0
HDBresale$LONGITUDE <- 0

for (i in 1:nrow(HDBresale)){
  temp_output <- geocode(HDBresale[i, 4], HDBresale[i, 5])
  
  HDBresale$LATITUDE[i] <- temp_output$results.LATITUDE
  HDBresale$LONGITUDE[i] <- temp_output$results.LONGITUDE
}
```

Converting HDBresale to an sf object

```{r}
HDBresale_sf <- HDBresale %>%
  st_as_sf(
    coords = c("LONGITUDE", "LATITUDE"),
    crs=4326) %>%
  st_transform(crs = 3414)
```

Import Preschool Location data

```{r}
preschool = st_read(
  "data/aspatial/PreSchoolsLocation.kml") %>%
  st_transform(crs= 3414)
```

Counting number of preschools

```{r}
prischool_within_1000m <- st_is_within_distance(
  HDBresale_sf, preschool, dist = 1000)

HDBresale_sf <- HDBresale_sf %>%
  mutate(prischool_within_1000m = map_int(prischool_within_1000m, length))
```

Import Childcare Location data

```{r}
childcare = st_read(
  "data/aspatial/ChildCareServices.kml") %>%
  st_transform(crs= 3414)
```

Counting number of childcare

```{r}
childcare_within_350m <- st_is_within_distance(
  HDBresale_sf, childcare, dist = 350)

HDBresale_sf <- HDBresale_sf %>%
  mutate(childcare_within_350m = map_int(childcare_within_350m, length))
```

Import kindergartens Location data

```{r}
kindergartens = st_read(
  "data/aspatial/Kindergartens.kml") %>%
  st_transform(crs= 3414)
```

Counting number of kindergarten

```{r}
kindergarten_within_350m <- st_is_within_distance(
  HDBresale_sf, kindergartens, dist = 350)

HDBresale_sf <- HDBresale_sf %>%
  mutate(kindergarten_within_350m = map_int(kindergarten_within_350m, length))
```

Import bus stop Location data

```{r}
busstop = st_read(dsn = "data/aspatial/BusStopLocation_Aug2025/",
                  layer = "BusStop") %>%
  st_transform(crs = 3414)
```

Counting number of busstop

```{r}
busstop_within_350m <- st_is_within_distance(
  HDBresale_sf, busstop, dist = 350)

HDBresale_sf <- HDBresale_sf %>%
  mutate(busstop_within_350m = map_int(busstop_within_350m, length))
```

Import eldercare service Location data

```{r}
eldercare = st_read(
  "data/aspatial/EldercareServicesKML.kml") %>%
  st_transform(crs= 3414)
```

compute proximity to eldercare service

```{r}
HDBresale_sf <- HDBresale_sf %>%
  mutate(
    PROX_eldercare = as.numeric(
      st_distance(geometry,
                  eldercare[st_nearest_feature(
                    geometry, eldercare), ], 
                  by_element = TRUE)
      )
    )
```

Import lrt mrt Location data

```{r}
lrtmrt = st_read(
  "data/aspatial/LTAMRTStationExitKML.kml") %>%
  st_transform(crs= 3414)
```

compute proximity to lrt and mrt

```{r}
HDBresale_sf <- HDBresale_sf %>%
  mutate(
    PROX_lrtmrt = as.numeric(
      st_distance(geometry,
                  lrtmrt[st_nearest_feature(
                    geometry, lrtmrt), ], 
                  by_element = TRUE)
      )
    )
```

Import hawker centre Location data

```{r}
hawker = st_read(
  "data/aspatial/HawkerCentresKML.kml") %>%
  st_transform(crs= 3414)
```

compute proximity to hawker centre

```{r}
HDBresale_sf <- HDBresale_sf %>%
  mutate(
    PROX_hawker = as.numeric(
      st_distance(geometry,
                  hawker[st_nearest_feature(
                    geometry, hawker), ], 
                  by_element = TRUE)
      )
    )
```

Import market and food centre Location data

```{r}
marketfoodcentre = st_read(
  "data/aspatial/NEAMarketandFoodCentreKML.kml") %>%
  st_transform(crs= 3414)
```

compute proximity to market and food centre

```{r}
HDBresale_sf <- HDBresale_sf %>%
  mutate(
    PROX_marketfoodcentre = as.numeric(
      st_distance(geometry,
                  marketfoodcentre[st_nearest_feature(
                    geometry, marketfoodcentre), ], 
                  by_element = TRUE)
      )
    )
```

Import supermarket location data

```{r}
supermarket <- read_csv("data/aspatial/ListofSupermarketLicences.csv")
```

Geocode supermarket postal codes

```{r}
postcodes <- unique(supermarket$postal_code)

url <- "https://onemap.gov.sg/api/common/elastic/search"

found <- data.frame()
not_found <- data.frame(postcode = character())

for (pc in postcodes) {
  query <- list(
    searchVal = pc,
    returnGeom = "Y",
    getAddrDetails = "Y",
    pageNum = "1"
  )
  
  res <- GET(url, query = query)
  json <- content(res)
  
  if (json$found != 0) {
    df <- as.data.frame(json$results, stringsAsFactors = FALSE)
    df$input_postcode <- pc
    found <- bind_rows(found, df)
  } else {
    not_found <- bind_rows(not_found, data.frame(postcode = pc))
  }
}
```

Appending location information to supermarket

```{r}
supermarket = supermarket %>%
  left_join(found, 
            by = c('postal_code' = 'POSTAL'))
```

Convert supermarket to sf object

```{r}
supermarket_clean <- supermarket %>%
  filter(!is.na(X) & !is.na(Y))

supermarket_sf <- st_as_sf(supermarket_clean, 
                         coords = c("X","Y"),
                         crs=3414) 
```

Compute proximity to supermarket

```{r}
HDBresale_sf <- HDBresale_sf %>%
  mutate(
    PROX_supermarket = as.numeric(
      st_distance(geometry,
                  supermarket_sf[st_nearest_feature(geometry, supermarket_sf), ],
                  by_element = TRUE)
    )
  )
```

Import park location data

```{r}
park <- st_read("data/aspatial/ParkFacilities.geojson") %>%
  st_transform(crs = 3414) 
```

Compute proximity to park

```{r}
HDBresale_sf <- HDBresale_sf %>%
  mutate(
    PROX_Park = as.numeric(
      st_distance(geometry, 
                  park[st_nearest_feature(geometry, park), ],
                  by_element = TRUE)
  ))
```

Using raffles place mrt coordinates as cbd

```{r}
cbd <- st_point(c(103.851959, 1.285400)) 

cbd_sf <- st_sfc(cbd, crs = 4326) %>%
  st_transform(crs = 3414)
```

Compute proximity to cbd

```{r}
HDBresale_sf <- HDBresale_sf %>%
  mutate(
    PROX_CBD = as.numeric(st_distance(geometry, cbd_sf))
  )
```

Compute age of unit

```{r}
HDBresale_sf <- HDBresale_sf %>%
  mutate(age_of_unit = 2025 - lease_commence_date)
```

Compute remaining lease

```{r}
HDBresale_sf <- HDBresale_sf %>%
  mutate(remaininglease = 99 - age_of_unit)
```

Changing storey to continuous value

```{r}
storey_map <- c(
  "01 TO 03" = 1,
  "04 TO 06" = 2,
  "07 TO 09" = 3,
  "10 TO 12" = 4,
  "13 TO 15" = 5,
  "16 TO 18" = 6,
  "19 TO 21" = 7,
  "22 TO 24" = 8,
  "25 TO 27" = 9,
  "28 TO 30" = 10,
  "31 TO 33" = 11,
  "34 TO 36" = 12,
  "37 TO 39" = 13,
  "40 TO 42" = 14,
  "43 TO 45" = 15,
  "46 TO 48" = 16,
  "49 TO 51" = 17
)

HDBresale_sf <- HDBresale_sf %>%
  mutate(floor_level = storey_map[storey_range])
```






