---
title: "In-class_Ex09"
author: "CHJ1"
---

Modelling Geographical Accessibility

Getting Started

```{r}
pacman::p_load(tmap, SpatialAcc, sf, 
               ggstatsplot, reshape2,
               tidyverse)
```

Importing geospatial data

```{r}
mpsz <- st_read(dsn = "data/geospatial", layer = "MP14_SUBZONE_NO_SEA_PL")
```

```{r}
hexagons <- st_read(dsn = "data/geospatial", layer = "hexagons") 
```

```{r}
eldercare <- st_read(dsn = "data/geospatial", layer = "ELDERCARE") 
```

Updating CRS information

```{r}
mpsz <- st_transform(mpsz, 3414)
eldercare <- st_transform(eldercare, 3414)
hexagons <- st_transform(hexagons, 3414)
```

```{r}
st_crs(mpsz)
```

Cleaning and updating attribute fields of the geospatial data

```{r}
eldercare <- eldercare %>%
  select(fid, ADDRESSPOS) %>%
  mutate(capacity = 100)
```

```{r}
hexagons <- hexagons %>%
  select(fid) %>%
  mutate(demand = 100)
```

Importing Distance Matrix

```{r}
ODMatrix <- read_csv("data/aspatial/OD_Matrix.csv", skip = 0)
```

Tidying distance matrix

```{r}
distmat <- ODMatrix %>%
  select(origin_id, destination_id, total_cost) %>%
  spread(destination_id, total_cost)%>%
  select(c(-c('origin_id')))
```

```{r}
distmat_km <- as.matrix(distmat/1000)
```

Modelling and Visualising Accessibility using Hansen Method

Computing Hansen’s accessibility

```{r}
acc_Hansen <- data.frame(ac(hexagons$demand,
                            eldercare$capacity,
                            distmat_km, 
                            #d0 = 50,
                            power = 2, 
                            family = "Hansen"))
```

```{r}
colnames(acc_Hansen) <- "accHansen"
```

```{r}
acc_Hansen <- as_tibble(acc_Hansen)
```

```{r}
hexagon_Hansen <- bind_cols(hexagons, acc_Hansen)
```

```{r}
acc_Hansen <- data.frame(ac(hexagons$demand,
                            eldercare$capacity,
                            distmat_km, 
                            #d0 = 50,
                            power = 0.5, 
                            family = "Hansen"))

colnames(acc_Hansen) <- "accHansen"
acc_Hansen <- as_tibble(acc_Hansen)
hexagon_Hansen <- bind_cols(hexagons, acc_Hansen)
```

Visualising Hansen’s accessibility

```{r}
mapex <- st_bbox(hexagons)
```

```{r}
tm_shape(hexagon_Hansen,
         bbox = mapex) + 
  tm_polygons(fill = "accHansen", 
              fill.scale = tm_scale_intervals(
                style = "quantile",
                n = 10,
                values = "brewer.blues")) +
tm_shape(eldercare) +
  tm_dots(size = 0.3,
          fill = "red",
          col = "black") +
  tm_title("Accessibility to eldercare: Hansen method") +
  tm_layout(frame = TRUE) +
  tm_compass(type="8star", size = 2) +
  tm_scalebar() +
  tm_grid(alpha = 0.2)
```

```{r}
powers <- seq(0.5, 2.5, 0.5)

acc_list <- list()

for (p in powers) {
  acc <- data.frame(
    accHansen = ac(hexagons$demand, eldercare$capacity, distmat_km,
                   power = p, family = "Hansen")
  ) %>%
    as_tibble() %>%
    mutate(power = p, fid = hexagons$fid)
  acc_list[[as.character(p)]] <- acc
}

acc_all <- bind_rows(acc_list)

hexagons_long <- left_join(
  expand.grid(fid = hexagons$fid, power = powers) %>%
    as_tibble(),
  acc_all, by = c("fid", "power")
)
hexagons_long <- left_join(hexagons_long, hexagons, by = "fid")
hexagons_long_sf <- st_as_sf(hexagons_long)
```

```{r}
tm_shape(hexagons_long_sf, bbox = mapex) +
  tm_polygons(
    col = "accHansen",
    palette = "Blues",
    style = "quantile",
    n = 10,
    title = "Accessibility"
  ) +
  tm_facets(by = "power", ncol = 2, free.coords = FALSE) +
  tm_shape(eldercare) +
  tm_dots(size = 0.3, fill = "red", col = "black") +
  tm_title("Accessibility to eldercare: Hansen method (powers 0.5 to 2.5)") +
  tm_layout(frame = TRUE) +
  tm_compass(type = "8star", size = 2) +
  tm_scalebar() +
  tm_grid(alpha = 0.2)
```

As Hansen power increases from 0.5 to 2.5, accessibility patterns become more clustered near elder care centres, with higher values decaying faster with distance and peripheral accessibility declining.

