---
title: "Take-home Exercise 1"
author: "CHJ1"
date: "September 13, 2025"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

In the first part of this assignment, I will be explaining each step and code chunk used for future reproducibility purposes.

Loads packages.

```{r}
pacman::p_load(tidyverse, sf, tmap, httr, rvest, spatstat, imager)
```

Lists and imports all CSV files whose names start with "ACRA". It combines the contents of these matched files into a single data frame called acra_data.

```{r}
folder_path <- "data"
file_list <- list.files(path = folder_path,
                        pattern = "^ACRA*.*\\.csv$",
                        full.names = TRUE)

acra_data <- file_list %>%
  map_dfr(read_csv)
```

Saves the acra_data data frame as an RDS file for later use, but is set not to run (eval: false) or print output (echo: false) in the document. This preserves data for quick loading in future analyses.

```{r}
#|eval: false
#|echo: false
write_rds(acra_data,
          "data/acra_data.rds")
```

Filters the imported dataset to retain selected columns and only businesses with specific SSIC codes, standardizes the date and postal code formats, creates year and month variables, and restricts output to records registered in 2025.

```{r}
biz_56111_55101_56130 <- acra_data %>%
  select(1:24) %>%
  filter(primary_ssic_code %in% c(56111, 55101, 56130)) %>%
  rename(date = registration_incorporation_date) %>%
  mutate(date = as.Date(date),
         YEAR = year(date),
         MONTH_NUM = month(date),
         MONTH_ABBR = month(date,
                      label = TRUE,
                      abbr = TRUE)) %>%
  mutate( 
    postal_code = str_pad(postal_code,
    width = 6, side = "left", pad = "0")) %>%
           filter(YEAR == 2025)
```

This code chunk performs postal code validation and enrichment by querying the Singapore OneMap API for each unique postal code in the filtered business dataset. It collects geospatial and address details for valid postal codes in the found data frame, while noting postal codes with no matches in not_found. This step enhances the dataset with location data for spatial analysis or mapping.

```{r}
postcodes <- unique(biz_56111_55101_56130$postal_code)

url <- "https://onemap.gov.sg/api/common/elastic/search"

found <- data.frame()
not_found <- data.frame(postcode = character())

for (pc in postcodes) {
  query <- list(
    searchVal = pc,
    returnGeom = "Y",
    getAddrDetails = "Y",
    pageNum = "1"
  )
  
  res <- GET(url, query = query)
  json <- content(res)
  
  if (json$found != 0) {
    df <- as.data.frame(json$results, stringsAsFactors = FALSE)
    df$input_postcode <- pc
    found <- bind_rows(found, df)
  } else {
    not_found <- bind_rows(not_found, data.frame(postcode = pc))
  }
}    
```

Reduces the dataset to the first 10 columns.

```{r}
found <- found %>%
  select(1:10)
```

Performs a left join to merge the business dataset biz_56111_55101_56130 with the postal code details stored in found, matching on the postal code columns from both datasets. This appends the geospatial and address information obtained from the OneMap API to each business record while preserving all businesses in the original dataset.

```{r}
biz_56111_55101_56130 = biz_56111_55101_56130 %>%
  left_join(found,
            by = c('postal_code' = 'POSTAL'))
```

Saves biz_56111_55101_56130 into an RDS file.

```{r}
write_rds(biz_56111_55101_56130, "data/biz_56111_55101_56130.rds")
```

Converts the biz_56111_55101_56130 data frame into a spatial (sf) object using the X and Y columns as coordinates and assigns the coordinate reference system EPSG:3414 for accurate spatial representation within Singapore’s national grid system.

```{r}
biz_56111_55101_56130_sf <- st_as_sf(biz_56111_55101_56130, 
                         coords = c("X","Y"),
                         crs=3414) 
```

Creates a bar plot to visualize the count of businesses by month. It provides a quick summary of trends in business registrations.

```{r}
ggplot(data = biz_56111_55101_56130,
       aes(x = MONTH_ABBR)) +
  geom_bar()
```

This code chunk switches the tmap visualization mode to interactive "view" mode and plots the spatial business points from biz_56111_55101_56130_sf as dots on an interactive map. This allows panning, zooming, and exploring the geographic distribution of businesses visually.

```{r}
tmap_mode('view')
tm_shape(biz_56111_55101_56130_sf)+
  tm_dots()
```

Filters the business data to include only records from the first half of 2025.

```{r}
biz_56111_55101_56130_sf_firsthalf2025 <- biz_56111_55101_56130_sf %>%
  filter(YEAR == 2025, MONTH_NUM %in% 1:6)
```

Splits the filtered spatial business dataset into separate data frames by the primary_ssic_code.

```{r}
split_list <- split(biz_56111_55101_56130_sf_firsthalf2025, 
                    biz_56111_55101_56130_sf_firsthalf2025$primary_ssic_code)

biz_56111_sf_firsthalf2025 <- split_list[["56111"]]
biz_55101_sf_firsthalf2025 <- split_list[["55101"]]
biz_56130_sf_firsthalf2025 <- split_list[["56130"]]
```

This code chunk reads a spatial polygon layer named "MP14_SUBZONE_WEB_PL" into mpsz.

```{r}
mpsz = st_read(dsn = "data", 
                  layer = "MP14_SUBZONE_WEB_PL")
```

This code chunk produces basic plots of the spatial polygons stored in the mpsz sf object.

```{r}
plot(mpsz)
```

This code chunk first queries the current CRS of mpsz. Then it explicitly sets the CRS of mpsz to EPSG 3414, ensuring the spatial data uses the correct projection for subsequent spatial operations and alignment with other datasets in the same CRS.

```{r}
st_crs(mpsz)
mpsz <- st_set_crs(mpsz, 3414)
```

Subsequent code chunks produce a layered plot where the base map of spatial polygons (mpsz) is drawn first, and the spatial points are overlaid on top, allowing visual examination of business locations within defined spatial boundaries.

```{r}
plot(st_geometry(mpsz))
plot(st_geometry(biz_56111_sf_firsthalf2025), add = TRUE)
```

```{r}
plot(st_geometry(mpsz))
plot(st_geometry(biz_55101_sf_firsthalf2025), add = TRUE)
```

```{r}
plot(st_geometry(mpsz))
plot(st_geometry(biz_56130_sf_firsthalf2025), add = TRUE)
```

This code chunk calculates the count of business points within each spatial polygon in mpsz by using the spatial intersection function st_intersects(). It stores this count in a new column. The following summary and top_n calls provide a quick view of the distribution of businesses per polygon and identify the polygon with the highest number of intersecting businesses.

```{r}
mpsz$`biz_56111_sf_firsthalf2025`<- lengths(st_intersects(mpsz, biz_56111_sf_firsthalf2025))
summary(mpsz$`biz_56111_sf_firsthalf2025`)

top_n(mpsz, 1, `biz_56111_sf_firsthalf2025`)
```

This code chunk calculates the area of each polygon feature in the mpsz spatial dataset using the st_area() function. It stores the resulting polygon areas in a new column called Area.

```{r}
mpsz$Area <- mpsz %>%
  st_area()
```

This code chunk calculates a new variable representing the business point density per square kilometer within each polygon by dividing the count by the polygon area and multiplying by 1,000,000 . It then produces a histogram to visualize the distribution of these density values across the polygons, providing insights into spatial concentration patterns of businesses.

```{r}
mpsz <- mpsz %>%
  mutate(`biz_56111_sf_firsthalf2025 Density` = `biz_56111_sf_firsthalf2025`/Area * 1000000)

hist(mpsz$`biz_56111_sf_firsthalf2025 Density`)
```

This code chunk creates a scatter plot, showing the relationship between the business counts and the business density per square km for each polygon in the mpsz dataset. The plot restricts axis limits to 0-40 for clearer visualization. This helps identify patterns or correlations between absolute counts and density values visually.

```{r}
ggplot(data=mpsz, 
       aes(y = `biz_56111_sf_firsthalf2025`, 
           x= as.numeric(`biz_56111_sf_firsthalf2025 Density`)))+
  geom_point(color="black", 
             fill="light blue") +
  xlim(0, 40) +
  ylim(0, 40) +
  labs(title = "",
      x = "biz_56111_sf_firsthalf2025 density (per km sq)",
      y = "biz_56111_sf_firsthalf2025 count")
```

```{r}
mpsz$`biz_55101_sf_firsthalf2025`<- lengths(st_intersects(mpsz, biz_55101_sf_firsthalf2025))
summary(mpsz$`biz_55101_sf_firsthalf2025`)

top_n(mpsz, 1, `biz_55101_sf_firsthalf2025`)
```

```{r}
mpsz <- mpsz %>%
  mutate(`biz_55101_sf_firsthalf2025 Density` = `biz_55101_sf_firsthalf2025`/Area * 1000000)

hist(mpsz$`biz_55101_sf_firsthalf2025 Density`)
```

```{r}
ggplot(data=mpsz, 
       aes(y = `biz_55101_sf_firsthalf2025`, 
           x= as.numeric(`biz_55101_sf_firsthalf2025 Density`)))+
  geom_point(color="black", 
             fill="light blue") +
  xlim(0, 40) +
  ylim(0, 40) +
  labs(title = "",
      x = "biz_55101_sf_firsthalf2025 density (per km sq)",
      y = "biz_55101_sf_firsthalf2025 count")
```

```{r}
mpsz$`biz_56130_sf_firsthalf2025`<- lengths(st_intersects(mpsz, biz_56130_sf_firsthalf2025))
summary(mpsz$`biz_56130_sf_firsthalf2025`)

top_n(mpsz, 1, `biz_56130_sf_firsthalf2025`)
```

```{r}
mpsz <- mpsz %>%
  mutate(`biz_56130_sf_firsthalf2025 Density` = `biz_56130_sf_firsthalf2025`/Area * 1000000)

hist(mpsz$`biz_56130_sf_firsthalf2025 Density`)
```

```{r}
ggplot(data=mpsz, 
       aes(y = `biz_56130_sf_firsthalf2025`, 
           x= as.numeric(`biz_56130_sf_firsthalf2025 Density`)))+
  geom_point(color="black", 
             fill="light blue") +
  xlim(0, 40) +
  ylim(0, 40) +
  labs(title = "",
      x = "biz_56130_sf_firsthalf2025 density (per km sq)",
      y = "biz_56130_sf_firsthalf2025 count")
```

This R function extract_kml_field extracts the value of a specific field from HTML table content provided as a string. It uses the rvest package to parse the HTML, find table rows (<tr>), and locate the row by matching the table header (<th>) text to the desired field name. Then it extracts the corresponding data cell (<td>) text, returning NA if the field or table is absent. This function is useful for scraping specific tabular data fields embedded within HTML documents or API responses.

```{r}
extract_kml_field <- function(html_text, field_name) {
  if (is.na(html_text) || html_text == "") return(NA_character_)
  
  page <- read_html(html_text)
  rows <- page %>% html_elements("tr")
  
  value <- rows %>%
    keep(~ html_text2(html_element(.x, "th")) == field_name) %>%
    html_element("td") %>%
    html_text2()
  
  if (length(value) == 0) NA_character_ else value
}
```

This code chunk reads a KML file "MasterPlan2019SubzoneBoundaryNoSeaKML.kml" into an sf object named mpsz2. It removes any Z and M dimensions from the geometries using st_zm(drop = TRUE, what = "ZM") and then transforms the coordinate reference system to EPSG 3414 for consistent spatial operations within Singapore's coordinate system. This prepares the subzone boundary spatial data for mapping or spatial analysis.

```{r}
mpsz2 <- st_read("data/MasterPlan2019SubzoneBoundaryNoSeaKML.kml") %>% 
  st_zm(drop = TRUE, what = "ZM") %>% st_transform(crs = 3414)
```

This code chunk extracts specific attributes ("REGION_N", "PLN_AREA_N", "SUBZONE_N", "SUBZONE_C") from the HTML in the Description using the extract_kml_field function, creating new columns with those extracted values. Then it removes the original Name and Description columns and relocates the geometry column to the end for cleaning and better data presentation.

```{r}
mpsz2 <- mpsz2 %>%
  mutate(
    REGION_N = map_chr(Description, extract_kml_field, "REGION_N"),
    PLN_AREA_N = map_chr(Description, extract_kml_field, "PLN_AREA_N"),
    SUBZONE_N = map_chr(Description, extract_kml_field, "SUBZONE_N"),
    SUBZONE_C = map_chr(Description, extract_kml_field, "SUBZONE_C")
  ) %>%
  select(-Name, -Description) %>%
  relocate(geometry, .after = last_col())
```

This code chunk filters the mpsz2 dataset to exclude subzones and planning areas related to offshore islands: "SOUTHERN GROUP" subzone, "WESTERN ISLANDS" and "NORTH-EASTERN ISLANDS" planning areas. 

```{r}
mpsz_cl <- mpsz2 %>%
  filter(SUBZONE_N != "SOUTHERN GROUP",
         PLN_AREA_N != "WESTERN ISLANDS",
         PLN_AREA_N != "NORTH-EASTERN ISLANDS")
```

This code chunk saves the filtered spatial dataset mpsz_cl to an RDS file. 

```{r}
write_rds(mpsz_cl, 
          "data/mpsz_cl.rds")
```

This code converts the mpsz_cl sf spatial polygon object into an observation window object of class owin. The as.owin function allows defining the study area boundary.

```{r}
sg_owin <- as.owin(mpsz_cl)
```

The subsequent code chunks convert the subset spatial points into a ppp. Because ppp requires an observation window, the points are subsequently restricted to the study area defined by sg_owin. This creates a spatstat-compatible point pattern restricted to Singapore.

```{r}
biz_56111_sf_firsthalf2025_ppp <- as.ppp(biz_56111_sf_firsthalf2025)
biz_56111_sf_firsthalf2025SG_ppp = biz_56111_sf_firsthalf2025_ppp[sg_owin]
```

```{r}
biz_55101_sf_firsthalf2025_ppp <- as.ppp(biz_55101_sf_firsthalf2025)
biz_55101_sf_firsthalf2025SG_ppp = biz_55101_sf_firsthalf2025_ppp[sg_owin]
```

```{r}
biz_56130_sf_firsthalf2025_ppp <- as.ppp(biz_56130_sf_firsthalf2025)
biz_56130_sf_firsthalf2025SG_ppp = biz_56130_sf_firsthalf2025_ppp[sg_owin]
```

This code runs the Clark and Evans spatial point pattern test on the point pattern of the business subsets. The test evaluates whether the points exhibit clustering, complete spatial randomness, or regularity within the boundary window. The specified settings:

correction="none" means no edge correction is applied.

clipregion="sg_owin" restricts the analysis to the study area defined by Singapore’s subzone boundary window.

alternative="clustered" tests for clustering (aggregation) as the alternative hypothesis.

method="MonteCarlo" uses Monte Carlo simulation with nsim=99 simulations to derive the test statistic distribution for significance assessment.

The result provides a Monte Carlo p-value testing whether the pattern is significantly clustered against the null hypothesis of randomness, helping interpret the spatial distribution pattern of the business locations within Singapore.

```{r}
clarkevans.test(biz_56111_sf_firsthalf2025SG_ppp,
                correction="none",
                clipregion="sg_owin",
                alternative=c("clustered"),
                method="MonteCarlo",
                nsim=99)
```

```{r}
clarkevans.test(biz_55101_sf_firsthalf2025SG_ppp,
                correction="none",
                clipregion="sg_owin",
                alternative=c("clustered"),
                method="MonteCarlo",
                nsim=99)
```

```{r}
clarkevans.test(biz_56130_sf_firsthalf2025SG_ppp,
                correction="none",
                clipregion="sg_owin",
                alternative=c("clustered"),
                method="MonteCarlo",
                nsim=99)
```

This code rescales the spatial point pattern from meters to kilometers.

Then, it computes a KDE density() with:

sigma=bw.diggle which specifies the bandwidth estimated by Diggle's method (for smoothing).

edge=TRUE to apply edge correction.

kernel="gaussian" indicating a Gaussian kernel smoothing function.

Finally, it plots the KDE heatmap showing spatial density of points representing the business locations. This KDE visualizes business intensity over space.

```{r}
biz_56111_sf_firsthalf2025SG_ppp_km <- rescale.ppp(
  biz_56111_sf_firsthalf2025SG_ppp, 1000, "km")
  
kde_biz_56111_sf_firsthalf2025SG_ppp_km <- density(biz_56111_sf_firsthalf2025SG_ppp_km,
                              sigma=bw.diggle,
                              edge=TRUE,
                              kernel="gaussian")
  
plot(kde_biz_56111_sf_firsthalf2025SG_ppp_km)
```

```{r}
biz_55101_sf_firsthalf2025SG_ppp_km <- rescale.ppp(
  biz_55101_sf_firsthalf2025SG_ppp, 1000, "km")
  
kde_biz_55101_sf_firsthalf2025SG_ppp_km <- density(biz_55101_sf_firsthalf2025SG_ppp_km,
                              sigma=bw.diggle,
                              edge=TRUE,
                              kernel="gaussian")
  
plot(kde_biz_55101_sf_firsthalf2025SG_ppp_km)
```

```{r}
biz_56130_sf_firsthalf2025SG_ppp_km <- rescale.ppp(
  biz_56130_sf_firsthalf2025SG_ppp, 1000, "km")
  
kde_biz_56130_sf_firsthalf2025SG_ppp_km <- density(biz_56130_sf_firsthalf2025SG_ppp_km,
                              sigma=bw.diggle,
                              edge=TRUE,
                              kernel="gaussian")
  
plot(kde_biz_56130_sf_firsthalf2025SG_ppp_km)
```

This code chunk filters the spatial polygons dataset mpsz_cl to select only the subzones within the "DOWNTOWN CORE" planning area of Singapore.

The Downtown Core is the central business district of Singapore, containing high-density commercial and financial activity areas, including prominent neighborhoods like Raffles Place, Marina Bay, and City Hall. This filtering isolates the polygons relevant to spatial analysis focused on this key urban core region within Singapore's Master Plan 2019 boundaries.

```{r}
dc <- mpsz_cl %>%
  filter(PLN_AREA_N == "DOWNTOWN CORE")
```

mfrow=c(2,2) sets the plotting layout to a 2 by 2 grid. The subsequent plot() command for st_geometry(dc) will plot the polygons for the Downtown Core subzones into the first slot of this grid, with the title "DOWNTOWN CORE".

```{r}
par(mfrow=c(2,2))
plot(st_geometry(dc), main = "DOWNTOWN CORE")
```

This code converts the "DOWNTOWN CORE" polygons sf object dc into owin. This conversion is needed to define the spatial study area boundary for spatstat analyses focused on just the Downtown Core region.

```{r}
dc_owin = as.owin(dc)
```

The code biz_XXXXX_sf_firsthalf2025_ppp[dc_owin] subsets the spatial point pattern object biz_XXXXX_sf_firsthalf2025_ppp to only include points that fall within the Downtown Core boundary dc_owin.

This spatial cropping restricts the analysis to the Downtown Core region, extracting only the business points located inside this key central area.

```{r}
biz_56111_sf_firsthalf2025_ppp_dc = biz_56111_sf_firsthalf2025_ppp[dc_owin] 
biz_55101_sf_firsthalf2025_ppp_dc = biz_55101_sf_firsthalf2025_ppp[dc_owin] 
biz_56130_sf_firsthalf2025_ppp_dc = biz_56130_sf_firsthalf2025_ppp[dc_owin]
```

This code rescales the spatial point pattern objects for the business categories from meters to kilometers.

```{r}
biz_56111_sf_firsthalf2025_ppp_dc.km = rescale.ppp(biz_56111_sf_firsthalf2025_ppp_dc, 1000, "km")
biz_55101_sf_firsthalf2025_ppp_dc.km = rescale.ppp(biz_55101_sf_firsthalf2025_ppp_dc, 1000, "km")
biz_56130_sf_firsthalf2025_ppp_dc.km = rescale.ppp(biz_56130_sf_firsthalf2025_ppp_dc, 1000, "km")
```

This code sets the plotting layout to a 2 by 2 grid so that multiple plots can be displayed together. Then it plots three unmarked spatial point patterns for the Downtown Core business categories.

Each plot will show the spatial distribution of the respective business points within the Downtown Core boundary.

```{r}
par(mfrow=c(2,2))
plot(unmark(biz_56111_sf_firsthalf2025_ppp_dc.km), 
  main="56111")
plot(unmark(biz_55101_sf_firsthalf2025_ppp_dc.km), 
  main="55101")
plot(unmark(biz_56130_sf_firsthalf2025_ppp_dc.km), 
  main="56130")
```

The set.seed(1234) sets the seed of the random number generator to ensure reproducibility of any subsequent random operations. By setting a fixed seed value, random processes such as simulation, sampling, or Monte Carlo methods produce the same results every time the code is run.

```{r}
set.seed(1234)
```

This code performs an analysis using the G-function (nearest neighbor distance distribution function) on the spatial point pattern object representing business points in the Downtown Core.

Gest() computes the empirical G-function with the edge correction method "border" to compensate for boundary effects in the spatial window.

plot() then visualizes the G-function curve with the x-axis limited to 0-500 meters, showing the cumulative distribution of nearest neighbor distances among points.

The G-function measures the distribution of distances from each point to its closest neighbor, useful for detecting clustering or regularity. The observed curve can be compared to those of theoretical patterns for inference on point pattern structure within the Downtown Core.

```{r}
G_56111 = Gest(biz_56111_sf_firsthalf2025_ppp_dc, correction = "border")
plot(G_56111, xlim=c(0,500))
```

```{r}
G_55101 = Gest(biz_55101_sf_firsthalf2025_ppp_dc, correction = "border")
plot(G_55101, xlim=c(0,500))
```

```{r}
G_56130 = Gest(biz_56130_sf_firsthalf2025_ppp_dc, correction = "border")
plot(G_56130, xlim=c(0,500))
```

These commands generate Monte Carlo simulation envelopes for the G-function for the three Downtown Core business point patterns:

envelope() runs 999 simulations of a CSR (completely spatially random) point pattern within the observation window.

For each simulation, the G-function is computed.

The function then constructs upper and lower envelopes of these simulated G-functions.

The observed G-function from the actual data is compared to these envelopes to assess significant departures from CSR (such as clustering or regularity).

This approach provides a graphical Monte Carlo test of spatial randomness by showing the range of expected G-functions under CSR and highlighting where the observed pattern deviates beyond chance.

```{r}
G_56111.csr <- envelope(biz_56111_sf_firsthalf2025_ppp_dc, Gest, nsim = 999)
G_55101.csr <- envelope(biz_55101_sf_firsthalf2025_ppp_dc, Gest, nsim = 999)
G_56130.csr <- envelope(biz_56130_sf_firsthalf2025_ppp_dc, Gest, nsim = 999)
```

These commands plot the Monte Carlo simulation envelopes previously computed for the G-function of the three Downtown Core business point patterns.

The plot shows the observed G-function (black curve),

The grey shaded region represents the simulation envelopes derived from CSR simulations,

And a red dashed line usually shows the theoretical G-function under CSR.

If the observed curve lies outside the grey envelope at certain distances, it indicates significant departure from spatial randomness (e.g., clustering or inhibition) at those scales.

Plotting these envelopes helps visually assess how the spatial point pattern compares to a null random model, supporting inference on spatial arrangement and interactions of these business points in the Downtown Core region.

```{r}
plot(G_56111.csr)
plot(G_55101.csr)
plot(G_56130.csr)
```

In the second part of this assignment, I will be describing and discussing the analysis results.

I am using primary_ssic_code 56111 (restaurants), 55101 (hotels) and 56130 (pubs) in my analysis.

```{r}
kde_56111 <- load.image("images/kde_56111.png")
plot(kde_56111)
```

This kernel density estimate (KDE) heat map shows the spatial concentration of business entities classified under code 56111 in Singapore for the first half of 2025. 

The colors indicate density, with dark blue showing low concentrations and yellow-purple areas showing higher intensity. Most of Singapore shows sparse distributions, but a few hotspots emerge where business activity is significantly clustered. The highest density area is in the Central Region around the Downtown Core and Orchard districts, which appear in bright yellow shades, reflecting the strongest clustering of these businesses.

Given that the Downtown Core is the central business district of Singapore, containing high-density commercial and financial activity areas, higher number of new restaurants is expected to open here. However, this could also be an indicator of higher number of old restaurants closing down due to factors such as high rental, giving way to new entrants.

```{r}
kde_55101 <- load.image("images/kde_55101.png")
plot(kde_55101)
```

This KDE heat map visualizes the spatial density of business entities classified under code 55101 in Singapore during the first half of 2025. 

Areas in dark blue indicate low business concentrations, while warmer yellow and pink tones represent higher densities. Most of the island shows very scattered distributions; however, several noticeable hotspots occur in the south-central part of Singapore. The highest density is observed in the Downtown Core, particularly near Orchard and Marina Bay, shown in bright yellow. This suggests these central districts remain the main clustering points for this business type, reflecting their commercial significance.

As the heart of Singapore and a popular tourist destination, the Downtown Core is a prime location for establishing hotels.

```{r}
kde_56130 <- load.image("images/kde_56130.png")
plot(kde_56130)
```

This KDE heat map shows where business entities of code 56130 are most concentrated in Singapore for the first half of 2025. 

Most of the map is dark blue, indicating very low densities overall. There are only a few locations with moderate clustering, shown as small yellow or magenta hotspots. The area with the highest density is in the Downtown Core and its nearby central districts, which display distinct yellow marks. This implies that, for this business type, Singapore’s central business area remains the main location for clustering.

This indicates that the demand for pubs is low in Singapore generally.

I will be performing the G function test on these business entities at Downtown Core in the subsequent sections since this is the area with high clustering for all 3 business entities.

```{r}
G_56111 <- load.image("images/G_56111.png")
plot(G_56111)
```

This G function plot evaluates the 2nd order spatial point pattern for 56111 business entities. The solid black line shows the observed nearest neighbor distribution, while the dashed red line represents the theoretical Poisson (random) expectation. The observed curve lies consistently above the Poisson curve at shorter distances, indicating stronger clustering: many points have a neighbor closer than would occur by chance. As radius r increases, curves converge, showing randomness at larger scales. This pattern suggests these businesses are spatially aggregated rather than uniformly or randomly distributed in Downtown Core.

```{r}
G_55101 <- load.image("images/G_55101.png")
plot(G_55101)
```

This G function plot investigates the 2nd order spatial point pattern for 55101 business entities. The solid black line (observed) sharply deviates above the red dashed line (Poisson expectation) at lower radii, before plateauing at 1. This suggests strong clustering of entities: many businesses have a neighbor much closer than expected under spatial randomness. The step-like shape indicates multiple locations with tightly packed entities and abrupt gaps. Overall, 55101 businesses are not randomly distributed—they are aggregated into discrete clusters in Downtown Core.

```{r}
G_56130 <- load.image("images/G_56130.png")
plot(G_56130)
```

This G function analysis for 56130 business entities shows the observed curve (solid black) never sharply exceeds the theoretical Poisson expectation (dashed red). This indicates no significant clustering: the spatial pattern of these entities is essentially random, with little evidence of aggregation or tight groupings. The 56130 businesses are more evenly dispersed in Downtown Core.

```{r}
G_56111.csr <- load.image("images/G_56111.csr.png")
plot(G_56111.csr)
```

This G function plot compares the observed spatial point pattern for 56111 entities (solid black line) to the theoretical random (CSR, dashed red) distribution, with simulation envelopes (gray area) indicating the expected range under randomness. The observed line is consistently above both the theoretical curve and the envelope, indicating significant clustering: many entities are closer together than expected by chance. Since the observed values lie above the envelope throughout the range, this finding is statistically significant and suggests non-random spatial aggregation.

```{r}
G_55101.csr <- load.image("images/G_55101.csr.png")
plot(G_55101.csr)
```

This G function plot for 55101 entities compares the observed nearest-neighbor distribution (solid black line) with the theoretical random pattern (dashed red) and includes confidence envelopes (gray area). The observed line rises steeply and reaches higher values much earlier than the theoretical curve, and it remains well above the simulation envelope across nearly the full range of r. This pattern demonstrates statistically significant clustering: these businesses are found much closer together than expected under complete spatial randomness, and the clustering effect is strong and highly unlikely to be due to chance.

```{r}
G_56130.csr <- load.image("images/G_56130.csr.png")
plot(G_56130.csr)
```

This G function plot for 56130 entities shows the observed nearest-neighbor distribution (solid line) compared with the theoretical random curve (dashed red), with simulation envelopes (grey area). The observed curve remains essentially flat and close to zero, sitting entirely below the simulation envelope, while the theoretical curve rises smoothly. This pattern indicates significant inhibition or regularity: these entities are spaced farther apart than expected under complete spatial randomness, showing spatial repulsion rather than clustering. The spatial distribution is more uniform, with entities avoiding close proximity to one another.

